Fusing Filters with Integer Linear Programming

### 2014-04-15 2014betterfusionforfilters/section/03-SizeInference.tex
working on size inference

\begin{tabbing}
MMMMM       \= MM \= MMMMMM \= \kill
$rate$      \> @::=@ \> $k_n$               \> (rate variable)\\
            \> $~|$  \> $rate \times rate$  \> (cross product) \\
%            \> $~|$  \> $Filtered_n$        \> (output of @filter@s) \\
%            \> $~|$  \> $External_n$        \> (output of @external@s) \\
\\
$constraint$ \> @::=@ \> $rate = rate$                      \\
             \> $~|$  \> $rate \le rate$                    \\
             \> $~|$  \> $constraint \wedge constraint$     \\
             \> $~|$  \> $\forall k_n. constraint$          \\
             \> $~|$  \> $\exists k_n. constraint$          \\
\end{tabbing}

@@

\infer{\lambda scalars~arrays \to @let @ binds : \forall_{a \in arrays} k_a. C}
      { \{k_a | a \in arrays\} \vdash binds : C}

@@

\infer{\Gamma \vdash o = @fold@~f~n; binds : C}
      {\Gamma \vdash binds                 : C}

@@

\infer{\Gamma \vdash o = @map@_n~f~ns; binds : \forall k_o. \bigwedge_{n \in ns}\{k_o = k_n\} \wedge C}
      {k_o, \Gamma \vdash binds                 : C}

@@

\infer{\Gamma \vdash o = @filter@~f~n; binds : \exists k_o. k_o \le k_n \wedge C}
      {k_o, \Gamma \vdash binds                 : C}

@@

\infer{\Gamma \vdash o = @gather@~i~d; binds : \forall k_o. k_o = k_i \wedge C}
      {k_o, \Gamma \vdash binds                 : C}

@@

\infer{\Gamma \vdash o = @cross@~a~b; binds : \forall k_o. k_o = k_a \times k_b \wedge C}
      {k_o, \Gamma \vdash binds                 : C}

@@

\infer{\Gamma \vdash outs = @external@~ins; binds : \exists_{o \in outs} k_o. C}
      {\{k_o | o \in outs\} \cup \Gamma \vdash binds                 : C}

@@

The generated equalities are easily grouped into equivalence classes.
All combinators whose constraints are in the same equivalence class have the same iteration size, and may be fused together.
These constraints are also used to rule out any programs that require runtime checks on array lengths during the program.

>>>>

\begin{tabbing}
MMMMM       \= MM \= MMMMMMMMMMM \= \kill
$rate$      \> @::=@ \> $k_n$               \> (rate variable)\\
            \> $~|$  \> $rate \times rate$  \> (cross product) \\
%            \> $~|$  \> $Filtered_n$        \> (output of @filter@s) \\
%            \> $~|$  \> $External_n$        \> (output of @external@s) \\
\\
$constraint$ \> @::=@ \> $rate = rate$                          \> (equality constraint) \\
             \> $~|$  \> $constraint \wedge constraint$         \> (conjunction) \\
             \> $~|$  \> $\forall k_n.\ constraint$              \> (forall quantification) \\
             \> $~|$  \> $\exists k_n.\ constraint$              \> (exists with no bounds)\\
             \> $~|$  \> $\exists k_n \le rate.\ constraint$     \> (exists with upper bound)    \\
\\
$\tau$       \> @::=@ \> $rate$                                  \> (loop size of some rate) \\
             \> $~|$  \> @external@                              \> (external and unfusible) \\
\end{tabbing}

\[ binds \vDash constraint : (\tau_{name} \Rightarrow \tau) \]

@@

\infer{\lambda scalars~arrays \to @let @ binds : \forall_{a \in arrays} k_a. C : T}
      { binds \vDash C : T}

@@

\infer{o = @fold@~f~n; binds \vDash C : \tau_o = k_n, T}
      {binds                 \vDash C : T}

@@

\infer{o = @map@_n~f~ns; binds \vDash \forall k_o.\ \bigwedge_{n \in ns}\{k_o = k_n\} \wedge C : \tau_o = k_o, T}
      {binds                   \vDash C : T}

@@

\infer{o = @filter@~f~n; binds \vDash \exists k_o \le k_n.\ C : \tau_o = k_n, T}
      {binds                   \vDash C : T}
Note that for a filter, the size of the output array $k_o$ is some existential that is less than or equal to $k_n$, but the actual loop size of the \emph{combinator} is equal to $k_n$.
This is because, in order to produce the filtered output, all elements of the input $n$ must be considered.

@@

\infer{o = @gather@~i~d; binds \vDash \forall k_o.\ k_o = k_i \wedge C : \tau_o = k_i, T}
      {binds                   \vDash C : T}

@@

\infer{o = @cross@~a~b; binds \vDash \forall k_o.\ k_o = k_a \times k_b \wedge C : \tau_o = k_a \times k_b, T}
      {binds                  \vDash C : T}

@@

\infer{outs = @external@~ins; binds \vDash \exists_{o \in outs} k_o.\ C : \tau_{outs} = @external@, T}
      {binds                        \vDash C : T}

@@

Solving the generated equality constraints allows each rate to be grouped into equivalence classes.
All combinators whose constraints are in the same equivalence class have the same iteration size, and may be fused together.
These constraints are also used to rule out any programs that require runtime checks on array lengths during the program.
